Gravitational-wave sensitivity curves

### 2014-09-24 V2/ReplyToReferees.txt

2)	We agree with this comment and think that it's well worth emphasising this point. We have added the following passage to the end of sec 5.4...

"It should be noted, however, that the sensitivity curves plotted in Appendix A would change substantially for stochastic backgrounds. This is because searches for stochastic backgrounds utilise the cross correlation of the outputs of multiple detectors for detection (Allen and Romano, 1999; Maggiore, 2000), instead of the individual outputs. Therefore the sensitivity of a network of ground based detectors is much greater than that of any individual detector. The shape of the sensitivity curve for a stochastic background would also differ from the curves plotted here, for a discussion of this point see Thrane and Romano (2013)."

Other comments (in order of appearance in the text):
1)	We thank the referee for pointing out this interesting source of confusion. We have checked this, and it seems that it was originally "evolved LISA", but the alternative "evolving LISA" has gradually become more popular and is now the prefered option. We have found numerous examples of both acronyms being used; for two conflicting examples on the same website see https://www.elisascience.org/news/top-news/gravitationaluniverseselectedasl3 and https://www.elisascience.org/articles/elisa-consortium/home . We have changed the main body of the text to read "evolving" added the following footnote...

"The e in eLISA originally stood for evolved, however, the preference is now for evolving."

@@

3)	We have made this correction.

4)	We have made this correction.

5)	We have made this correction.

6)	We agree with the referee on this comment. The message we were trying to convey here is that current limits are comparable in order of magnitude. We have changed the relevant passage in sec 4.4.1 to now state that the PPTA limit is lower by a factor of ~2.5. Also see comment (7) for more changes to section 4.4.1.

7)	We agree with the referee's comment that more information on the assumptions behind the PTA sensitivity curves in required. We have modified sections 4.4.1, 4.4.2 and 4.4.3 so that they now clearly state the total time, frequency of measurement, r.m.s. timing precision and number of pulsars for each PTA.

### 2014-09-24 V2/Detectors.tex
done point 7

The PTAs currently in operation are the European Pulsar Timing Array \citep[EPTA\footnote{\url{http://www.epta.eu.org/}};][]{eptareview2013}, the Parkes Pulsar Timing Array \citep[PPTA\footnote{\url{http://www.atnf.csiro.au/research/pulsar/ppta/}};][]{parkesreview2013} based in Australia, and the North American Nanohertz Observatory for Gravitational waves \citep[NANOGrav\footnote{\url{http://nanograv.org/}};][]{nanogravreview2013}. There are published limits on the amplitude of the stochastic background from all three detectors: the most recent from EPTA is \citet{Haasteren}, from PPTA is \citet{Shannon2013} and from NANOGrav is \citet{2013ApJ...762...94D}. The limits from the existing detectors are all of a similar magnitude; the PPTA limit is a factor of $\sim 2.5$ lower than the other two. In \ref{app:a} we use the EPTA limits, this was based on an analysis of $5$ pulsars over approximately $10~\mathrm{yr}$. The curve labelled ``EPTA'' is normalised to agree with the limit published in \cite{Haasteren} at a frequency of $f_{0}=1~\mathrm{yr^{-1}}$.

@@

Combining the existing arrays would yield a single PTA using approximately $4$ times as many pulsars; this consortium of consortia is known as the International Pulsar Timing Array \citep[IPTA\footnote{\url{http://www.ipta4gw.org/}};][]{iptareview2013}. The IPTA curves plotted in the figures are based on $20$ pulsars timed for twice as long as the current EPTA.

@@

The next great advancement in radio astronomy shall come with the completion of the Square Kilometre Array \citep[SKA;][]{Dewdney2009}. This shall greatly increase the sensitivity of pulsar timing \citep{Kramer2004}. The sensitivity curves plotted in \ref{app:a} for the SKA assume a root-mean-square error on the timing residuals a factor of $10$ better than current PTAs, and include $20$ pulsars timed for twice as long as the current EPTA.

>>>>

The PTAs currently in operation are the European Pulsar Timing Array \citep[EPTA\footnote{\url{http://www.epta.eu.org/}};][]{eptareview2013}, the Parkes Pulsar Timing Array \citep[PPTA\footnote{\url{http://www.atnf.csiro.au/research/pulsar/ppta/}};][]{parkesreview2013} based in Australia, and the North American Nanohertz Observatory for Gravitational waves \citep[NANOGrav\footnote{\url{http://nanograv.org/}};][]{nanogravreview2013}. There are published limits on the amplitude of the stochastic background from all three detectors: the most recent from EPTA is \citet{Haasteren}, from PPTA is \citet{Shannon2013} and from NANOGrav is \citet{2013ApJ...762...94D}. The limits from the existing detectors are all of a similar magnitude. In \ref{app:a} we use the EPTA limit based on an analysis of $5$ pulsars over approximately $10~\mathrm{years}$. The curve labelled ``EPTA'' assumes that all the pulsar were timed identically every 2 weeks with an r.m.s.\ error in each timing residual of $100\,\textrm{ns}$. The NANOGrav limit is comparable to this, and the PPTA limit is a factor $\sim 2.5$ lower.

@@

Combining the existing arrays would yield a single PTA using approximately $4$ times as many pulsars; this consortium of consortia is known as the International Pulsar Timing Array \citep[IPTA\footnote{\url{http://www.ipta4gw.org/}};][]{iptareview2013}. The IPTA curve plotted in the figures assumes $20$ pulsars timed every 2 weeks for 15 years with an r.m.s.\ error in each timing residual of $100\,\textrm{ns}$.

@@

The next great advancement in radio astronomy shall come with the completion of the Square Kilometre Array \citep[SKA;][]{Dewdney2009}. This shall greatly increase the sensitivity of pulsar timing \citep{Kramer2004}. The sensitivity curves plotted in \ref{app:a} for the SKA assumes $50$ pulsars timed every 2 weeks for 15 years with an r.m.s.\ error in each timing residual of $100\,\textrm{ns}$. Our choice of 50 pulsars might be quite conservative, it is possible that the SKA will discover many more suitable milli-second pulsars than this but (initially at least) these won't all have been timed for at least a few years.
